---
title: "Week 3 & 4: Sentiment Analysis and TF-IDF"
author: "Viola Thadtanone"
format: html
editor: visual
---

## Setup & Preparation

### Import Library

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)
library(ggplot2)
library(forcats)
library(tibble)
library(scales)
library(quanteda)
library(quanteda.textstats)
```

### Import input text files

```{r}
# Read the raw text files into R
circle_raw <- read_file('texts/A07594__Circle_of_Commerce.txt')
free_raw <- read_file('texts/B14801__Free_Trade.txt')
txt_third <- read_file('texts/A06785.txt')

# Combine into a tibble for tidytext workflows
texts_miss <- tibble(
  doc_title = c("Circle of Commerce", "Free Trade"),
  text = c(circle_raw, free_raw)
)
texts_miss
```

## I. Raw-Count Sentiment Analysis

### Step 1: Tokenize and clean the text:

```{r}
# Perform data cleaning and tokenize overall texts including the assumptions that
# 1. Remove punctuation since it does not affect the meaning of the words
# 2. Remove numbers since it may include irrelevant numbers from pagination
# 3. Remove symboles since it is likely a transcription/OCR errors
tokens_1 <- texts_miss %>%
  mutate(
    text = text %>%
      str_replace_all("ſ", "s") %>%
      str_replace_all("\\s+", " ")
  ) %>%
  unnest_tokens(word, text) %>%          # This already removed punctuation
  mutate(word = str_to_lower(word)) %>%
  filter(
    !str_detect(word, "^[0-9]+$"),
    !str_detect(word, "[^[:alnum:]]")
  )

tokens_1
```

```{r}
# Start with tidytext's built-in stopword list - This is such as "the", "a", "an".
data("stop_words")

# Add our own project-specific stopwords (you can, and will, expand this list later)
custom_stopwords <- tibble(
  word = c(
    "vnto", "haue", "doo", "hath", "bee", "ye", "thee"
  )
)

all_stopwords <- bind_rows(stop_words, custom_stopwords) %>%
  distinct(word)

all_stopwords
```

### Remove stopwords

```{r}
tokens_1_clean <- tokens_1 %>%
  
  # Remove stopwords - Only keep the unmatch one
  anti_join(all_stopwords, by = "word")

tokens_1_clean
```

### Step 2: Join tokens to the Bing sentiment dictionary

```{r}
# get_sentiments("bing") returns a table with two columns: word and sentiment
bing <- get_sentiments("bing")
```

### Step 3: Compute raw sentiment totals

```{r}
# Create a summary table for raw sentiment totals
sentiment_summary1 <- tokens_1_clean %>%
  inner_join(bing, by = "word") %>%
  count(doc_title, sentiment) %>%
  pivot_wider(
    names_from = sentiment,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(net_sentiment = positive - negative)

sentiment_summary1
```

## II. TF-IDF–Weighted Sentiment Analysis

### Step 4: Compute TF-IDF for words in each document

```{r}
texts <- c(
  "Circle of Commerce" = circle_raw,
  "Free Trade"         = free_raw,
  "Malynes"  = txt_third
)

# Perform data cleaning similar to the previous analyis
texts <- texts %>%
  str_replace_all("ſ", "s") %>%
  str_replace_all("\\s+", " ") %>%
  str_to_lower()

# Incorporate into Corpus
corp <- corpus(texts)
corp
```

```{r}
# Perform tokenization with same assumptions with Raw-Count Sentiment Analysis
tokens_2_clean <- tokens(
  corp,
  remove_punct = TRUE,
  remove_numbers = TRUE,
  remove_symbols = TRUE
)

# Remove the same set of stopwords 
tokens_2_clean <- tokens_2_clean %>%
  tokens_remove(all_stopwords$word)

tokens_2_clean
```

```{r}
# Create Document-Feature Matrix using all 3 documents
dfm_mat <- dfm(tokens_2_clean)
dfm_mat
```

```{r}
# Calculate TF-IDF using the full corpus
dfm_tfidf_mat <- dfm_tfidf(dfm_mat)
dfm_tfidf_mat
```

### Step 5: Convert the TF-IDF results to a tidy table

```{r}
# Convert the matrix into a regular data frame
tfidf_df <- as.data.frame(as.matrix(dfm_tfidf_mat))
tfidf_df$doc_title <- rownames(tfidf_df)
tfidf_df
```

```{r}
# Convert to tiday format and join the sentimate from 
tfidf_tidy <- tfidf_df %>%
  pivot_longer(
    cols = -doc_title,
    names_to = "word",
    values_to = "tfidf"
  ) %>%
  filter(tfidf > 0) %>%  # Remove rows where the TF-IDF weight is 0
  filter(doc_title %in% c("Circle of Commerce", "Free Trade")) # Only kept the first 2 documents

tfidf_tidy
```

### Step 6: Keep only sentiment-bearing words

```{r}
# Join tidy table to the Bing sentiment dictionary

tfidf_bing <- tfidf_tidy %>%
  inner_join(bing, by = "word") %>%
  arrange(doc_title, desc(tfidf))

tfidf_bing
```

### Step 7: Compute TF-IDF–weighted sentiment totals

```{r}
# Create a summary table 
sentiment_summary2 <- tfidf_bing %>%
  group_by(doc_title, sentiment) %>%
  summarise(total_tfidf = sum(tfidf)) %>%
  pivot_wider(
    names_from = sentiment,
    values_from = total_tfidf,
    values_fill = 0  # Fill NA with 0
  ) %>%
  mutate(net_tfidf_sentiment = positive - negative)

sentiment_summary2
```

## III. Compare Raw vs. TF-IDF Sentiment

```{r}
# Create final table for csv export
final_table <- sentiment_summary1 %>%
  left_join(sentiment_summary2, by = "doc_title") %>%
  rename(
    positive_raw = positive.x,
    negative_raw = negative.x,
    net_sentiment_raw = net_sentiment,
    positive_tfidf = positive.y,
    negative_tfidf = negative.y
  )

final_table
```
